---
title: "FFMS: Flexible Frequentist Model Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FFMS-guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette introduces the **FFMS** package and shows how to perform **Flexible Frequentist Model Selection** for generalized nonlinear models using genetically modified feature search combined with a Continuous $\ell_0$ Relaxation known as the Smooth Information Criterion (SIC).

------------------------------------------------------------------------

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = TRUE,  # show package startup and other messages
  warning = FALSE, # suppress warnings
  echo    = TRUE,  # show code
  results = "markup"
)
```

```{r}
library(FFMS)
```

------------------------------------------------------------------------

# Example 1 — FFMS: recovering Kepler’s third law

**Data**: $n=612$ exoplanets; variables include `semimajoraxis`, `period`, `hoststar_mass`, etc.
Target relationship: ${a \approx K_2 \left(P^2 M_h\right)^{1/3}}$.

We define a set of transforms and run the basic single-threaded `ffms_base` method on a subset of the data. 

```{r}
# Load example dataset
data <- FFMS::exoplanet
df.train <- data[1:500, ]

# Define transforms
to3 <- function(x) x^3
p2  <- function(x) x^2
transforms <- c("sigmoid", "sin_deg", "exp_dbl", "p0", "p2", "troot", "to3")

# Fast single-thread run (set small for the vignette)
result <- ffms(
  formula     = semimajoraxis ~ 1 + .,
  data        = df.train,
  method      = "ffms_base",
  transforms  = transforms,
  pop.max     = 10,
  P           = 5,
  N           = 100
)

# Summarize the top models found
summary(result)
```

# Parallel Run 

For difficult problems where deep non-linear combinations are needed (e.g., $M^{1/3} \cdot P^{2/3}$ which requires chaining 3 transformations), we run `ffms.parallel`. This method runs multiple evolutionary chains independently and merges their discovered features.

```{r eval=FALSE}
# Example of a fully tuned parallel run that reliably finds Kepler's Third Law
result.parallel <- ffms(
  formula     = semimajoraxis ~ 1 + .,
  data        = df.train,
  method      = "ffms.parallel",
  transforms  = transforms,
  runs        = 16,
  pop.max     = 80,
  prob_gen    = c(0.5, 0.3, 0.1, 0.1), # favour multiplication for chaining
  prob_filter = 0.3,
  penalty_a   = 2.0,                   # heavy penalty for complexity
  P           = 100,
  N           = 500,
  cores       = parallel::detectCores() - 1
)

summary(result.parallel)
```
